Datasheet for NBA player stats

### 2024-04-08 paper/datasheet_template.qmd
test file finished

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

>>>>

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was directly observable and publicly available, compiled from official NBA games and player performance statistics. These statistics are typically recorded by official scorers and statisticians during games and are not reported by the subjects (players) themselves. The data is validated through the official NBA data validation processes, which involve real-time and post-game reviews to ensure accuracy.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The collection involved both hardware apparatuses (like tracking devices and cameras in arenas) and software programs (for data aggregation and presentation). NBA and SportRadar, as the official statistics provider, likely use a combination of these tools to compile and validate game statistics. The validation processes are rigorous, involving multiple checks to ensure data accuracy and reliability.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- This dataset is not a sample but rather a comprehensive compilation of all player statistics for the 2023-2024 NBA season, representing a complete set of data points for that timeframe.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data was collected by official NBA statisticians and game officials, possibly with support from SportRadar personnel. These individuals are professionals employed or contracted by the NBA and SportRadar, compensated according to their professional agreements, not crowdworkers.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected over the course of the 2023-2024 NBA season. The timeframe of data collection matches the creation timeframe of the data, as it is recorded live during games and updated in near real-time.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Given the public and professional nature of the data, formal ethical review processes like those conducted by institutional review boards are generally not applicable. The data collection is part of standard operations within professional sports leagues.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was obtained via a third party, the Basketball Reference website, which aggregates official NBA statistics.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- As professional athletes, NBA players are aware that their performance metrics are recorded and publicly disseminated as part of their participation in the league. Formal individual notifications for each data collection instance are not standard practice.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Players consent to the recording and public sharing of their game performance as part of their professional contracts with the NBA and their respective teams. Explicit consent for each instance of data collection in the form of game statistics is inherent to their professional engagement.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- The public sharing of game performance data is a standard part of professional sports contracts. Individual players typically do not have the option to revoke consent for the sharing of standard game performance data.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- An analysis of this nature is generally not conducted for publicly available professional sports performance data, as it is considered non-sensitive and is a standard aspect of professional sports analytics.
12. *Any other comments?*
	- The collection and use of this data are standard practices within the domain of professional sports analytics and are not subject to the typical privacy concerns associated with personal or sensitive data.
